---
title: "Data Analysis Project Proposal"
author: "Course: STAT 420 (Summer 2021), Authors: Hugh Huang, Manupriya Arora, Son-Tung Nguyen"
date: ''
output:
  html_document: 
    theme: readable
    toc: yes
urlcolor: cyan
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Analysis of AirBnB Listing Prices in Vancouver, BC, CA

## Data loading & cleaning

```{r, message = FALSE}
# Load dataset
library(readr)
listings = read_csv("vancouver_listings_cleansed.csv")

# Convert some columns to factors
listings$host_neighbourhood = as.factor(listings$host_neighbourhood)
listings$neighbourhood_cleansed = as.factor(listings$neighbourhood_cleansed)
listings$property_type = as.factor(listings$property_type)
listings$room_type = as.factor(listings$room_type)

# Remove id and string columns
listings$id = NULL
listings$description = NULL
listings$host_acceptance_rate = NULL
listings$host_verifications = NULL
listings$amenities = NULL
listings$host_id = NULL
listings = na.omit(listings)

listings = subset(listings, listings$host_neighbourhood != "Kihei/Wailea")
listings = subset(listings, listings$host_neighbourhood != "Ewa")
listings = subset(listings, listings$property_type != "Camper/RV")
listings = subset(listings, listings$property_type != "Entire floor")
listings = subset(listings, listings$property_type != "Private room in boat")
listings = subset(listings, listings$property_type != "Shared room in hostel")
listings = subset(listings, listings$property_type != "Private room in bed and breakfast")
listings = subset(listings, listings$property_type != "Room in bed and breakfast")
listings = subset(listings, listings$host_neighbourhood != "Metrotown")
listings = subset(listings, listings$room_type != "Hotel Room")

#listings$price[which(listings$price > 1000)]
#listings = subset(listings, listings$price < 1200) #removing outliers
```

## Test:Train Split

```{r}
set.seed(420)
listings_trn_idx  = sample(nrow(listings), size = trunc(0.80 * nrow(listings)))
listings_trn_data = listings[listings_trn_idx, ]
listings_tst_data = listings[-listings_trn_idx, ]
```

## Model building

```{r}
full_add_model = lm(price ~ ., data = listings_trn_data)
#m_aic = step(full_add_model, direction = "both", trace = 0)
library(leaps)
head(listings_trn_data)

listings_trn_data_v1 = subset(listings_trn_data, select = -c(host_neighbourhood, host_listings_count, reviews_per_month, has_availability, availability_365, review_scores_cleanliness, property_type))

listings_trn_data_v2 = subset(listings_trn_data, select = -c(host_neighbourhood, host_listings_count, reviews_per_month, has_availability, availability_365, review_scores_cleanliness, neighbourhood_cleansed))

listings_trn_data_v3 = subset(listings_trn_data, select = -c(property_type, host_listings_count, reviews_per_month, has_availability, availability_365, review_scores_cleanliness, neighbourhood_cleansed))

sum(listings$room_type == "Hotel room")
listings$room_type = droplevels(listings$room_type, exclude = "Hotel room")
levels(listings$room_type)
listings_trn_data$room_type = droplevels(listings_trn_data$room_type, exclude = "Hotel room")
levels(listings_trn_data$room_type)

ex_v1 = summary(regsubsets(price ~ ., data = listings_trn_data_v1, really.big = TRUE))
ex_v1$which[9,]

ex_v2 = summary(regsubsets(price ~ ., data = listings_trn_data_v2, really.big = TRUE))
ex_v2$which[9,]

ex_v3 = summary(regsubsets(price ~ ., data = listings_trn_data_v3, really.big = TRUE))
ex_v3

ex_nc = lm(price ~ neighbourhood_cleansed + accommodates +bathrooms_cleansed + bedrooms + review_scores_rating, data = listings_trn_data) #no property_type or host_neighbourhood

ex_pt = lm(price ~ property_type + accommodates +bathrooms_cleansed + bedrooms + review_scores_rating, data = listings_trn_data) #no neighbourhood_cleansed or host_neighbourhood

ex_hn = lm(price ~ host_neighbourhood + accommodates +bathrooms_cleansed + bedrooms + review_scores_rating, data = listings_trn_data) #no property_type or neighbourhood_cleansed
```

```{r}
ex_try = lm(price ~ neighbourhood_cleansed + property_type + bathrooms_cleansed + bedrooms + neighbourhood_cleansed:bedrooms + accommodates + review_scores_rating, data = listings_trn_data)

ex_try_v2 = lm(price ~ neighbourhood_cleansed + property_type + bathrooms_cleansed + bedrooms + neighbourhood_cleansed:bedrooms + accommodates + review_scores_rating + bathrooms_cleansed:bedrooms + bathrooms_cleansed:accommodates + bathrooms_cleansed:review_scores_rating + bedrooms:accommodates + bedrooms:review_scores_rating + accommodates:review_scores_rating, data = listings_trn_data)

ex_try_v3 = lm(price ~ neighbourhood_cleansed + property_type + bathrooms_cleansed + bedrooms + neighbourhood_cleansed:bedrooms + accommodates + review_scores_rating + bathrooms_cleansed:bedrooms + bathrooms_cleansed:accommodates + bathrooms_cleansed:review_scores_rating + bedrooms:accommodates + bedrooms:review_scores_rating + accommodates:review_scores_rating + review_scores_cleanliness + has_availability + room_type, data = listings_trn_data)

#full_interaction_model = lm(price ~ . + ^ 2, data = listings_trn_data)

ex_v4 = summary(regsubsets(price ~ (bathrooms_cleansed + bedrooms + accommodates + review_scores_rating) ^ 2, data = listings_trn_data, really.big = TRUE))
ex_v4$which[8,]
```

```{r}
plot(price ~ ., data = listings_trn_data)
boxplot(listings$price)
```


##Diagnostics

```{r}
library(lmtest)
diagnostics = function(model, pcol = 'grey', lcol = 'dodgerblue', alpha = 0.05, plotit = TRUE, testit = TRUE) {
  if (plotit) {
    par(mfrow=c(1, 2))
    
    plot(fitted(model), resid(model), col = pcol, xlab = "Fitted", ylab = "Residuals", main = "Residuals vs fitted")
    abline(h = 0, col = lcol, lwd = 2)
    
    qqnorm(resid(model), main = "Normal Q-Q Plot", col = pcol)
    qqline(resid(model), col = lcol, lwd = 2)

  }
  
  if (testit) {
    shapiro_wilk = shapiro.test(resid(model))
    decision = if (shapiro_wilk$p.value < alpha) "Reject" else "Fail to Reject"
    list("p_val" = shapiro_wilk$p.value, "decision" = decision)
  }
}

get_loocv_rmse = function(model) {
  sqrt(mean((resid(model) / (1 - hatvalues(model))) ^ 2))
}

get_test_average_percent_error = function(model) {
  test_predictions = predict(model, newdata = listings_tst_data)
  sum(abs(test_predictions - listings_tst_data$price) / test_predictions) * 100 / length(listings_tst_data$price)
}

get_test_rmse = function(model){
  sqrt(mean((listings_tst_data$price - predict(model, listings_tst_data))^2))
}
```

```{r}
get_test_average_percent_error(ex_nc)
get_loocv_rmse(ex_nc)

get_test_average_percent_error(ex_pt)
get_loocv_rmse(ex_pt)

get_test_average_percent_error(ex_hn)
get_loocv_rmse(ex_hn)

get_test_average_percent_error(full_add_model)
get_loocv_rmse(full_add_model)

get_test_average_percent_error(ex_try_v2)
get_loocv_rmse(ex_try_v2)

get_test_average_percent_error(ex_try_v3)
get_loocv_rmse(ex_try_v3)

diagnostics(ex_try_v2)
```

#experimenting

ex_try_v3 = lm(price ~ neighbourhood_cleansed + property_type + bathrooms_cleansed + bedrooms + neighbourhood_cleansed:bedrooms + accommodates + review_scores_rating + bathrooms_cleansed:bedrooms + bathrooms_cleansed:accommodates + bathrooms_cleansed:review_scores_rating + bedrooms:accommodates + bedrooms:review_scores_rating + accommodates:review_scores_rating + review_scores_cleanliness + has_availability + room_type, data = listings_trn_data)

```{r}
summary(ex_try_v3)

exp = lm(price ~ accommodates + bathrooms_cleansed + bedrooms + log(review_scores_rating), data = listings_trn_data)

summary(exp)

get_test_average_percent_error(exp)
get_loocv_rmse(exp)
get_test_rmse(exp)

test_predictions = predict(exp, newdata = listings_tst_data)
which(abs(test_predictions - listings_tst_data$price) == max(abs(test_predictions - listings_tst_data$price)))

abs(test_predictions - listings_tst_data$price) / test_predictions

diagnostics(exp)
```

##experiment with exp but removing outliers
```{r}
sum(cooks.distance(exp) > 4 / length(cooks.distance(exp)))

cd_exp = cooks.distance(exp)
exp2 = lm(price ~ accommodates + bathrooms_cleansed + bedrooms + log(review_scores_rating), data = listings_trn_data, subset = cd_exp <= 4 / length(cooks.distance(exp)))

get_test_average_percent_error(exp2)
get_loocv_rmse(exp2)
get_test_rmse(exp2)

test_predictions2 = predict(exp2, newdata = listings_tst_data)
max(abs(test_predictions2 - listings_tst_data$price))

which(abs(test_predictions2 - listings_tst_data$price) == max(abs(test_predictions2 - listings_tst_data$price)))

listings_tst_data[134,]
test_predictions2[134]
```


##exp but adding terms
```{r}
exp_nc = lm(price ~ accommodates + bathrooms_cleansed + bedrooms + log(review_scores_rating) + neighbourhood_cleansed, data = listings_trn_data) # bad

summary(exp_nc)
get_test_average_percent_error(exp_nc)
get_loocv_rmse(exp_nc)
get_test_rmse(exp_nc)

#+ accommodates:bathrooms_cleansed + bathrooms_cleansed:bedrooms

exp_int = lm(price ~ accommodates + bathrooms_cleansed + bedrooms + log(review_scores_rating), data = listings_trn_data)

summary(exp_int)
get_test_average_percent_error(exp_int)
get_loocv_rmse(exp_int)
get_test_rmse(exp_int)
```


##experimenting with outliers, modify cooks threshold
```{r}
sum(cooks.distance(exp) > 10 / length(cooks.distance(exp)))

exp3 = lm(price ~ accommodates + bathrooms_cleansed + bedrooms + log(review_scores_rating), data = listings_trn_data, subset = cd_exp <= 10 / length(cooks.distance(exp)))

get_test_average_percent_error(exp3)
get_loocv_rmse(exp3)
get_test_rmse(exp3)
```

